---
title: "R Notebook"
output: html_notebook
---

# Input
```{r}
rw.merge <- read.csv("C:\\Users\\tenni\\OneDrive\\Documents\\GitHub\\r_w_project\\data\\rw_merge.csv")
acous <- read.csv("C:\\Users\\tenni\\OneDrive\\Documents\\GitHub\\r_w_project\\data\\acoustic.csv")
```

child age
```{r}
acous$ChildAge <- as.numeric(substr(acous$ChildID, start=5, stop=6))
rw.merge$ChildAge <- as.numeric(substr(rw.merge$ChildID, start=5, stop=6))
```

libraries
```{r, warning=FALSE, message=FALSE, echo=FALSE}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(ggpubr)
library(GGally)
```

# F2F3 for R and W

How to scale? 
- Account for age
- scale F3 F2 separately, then calculate their difference

How should we account for age? 

## f2 by age for /r/ and /w/
```{r}
ggplot(
  data=(acous %>% filter(
    tran == c("$r", "r", "$w", "w")
  )), 
  aes(y=f2, x=ChildAge, color=tran)
) + 
  scale_colour_manual(values=cbPalette) + theme_tufte()+
  geom_point()+
  geom_smooth(method=loess)
```

## f3 by age for /r/ and /w/
```{r}
ggplot(
  data=(acous %>% filter(
    tran == c("$r", "r", "$w", "w")
  )), 
  aes(y=f3, x=ChildAge, color=tran)
) + 
  scale_colour_manual(values=cbPalette) + theme_tufte()+
  geom_point()+
  geom_smooth(method=loess)
```
The f3 drops for r, its like a refinement. Or for younger kids there's a contributor, or a proxy, or another cue. 

Is the other cue vowels? 

### f3 by age for /r/ by vowel
```{r}
ggplot(
  data=(acous %>% filter(
    tran == c("r")
  )), 
  aes(y=f3, x=ChildAge, color=Letter)
) + 
  scale_colour_manual(values=cbPalette) + theme_tufte()+
  geom_point()+
  geom_smooth(method=loess, alpha=.1)
```
Nope.

## f3f2 difference by age for /r/
```{r}
ggplot(
  data=(acous %>% filter(
    tran == c("$r", "r", "$w", "w")
  )), 
  aes(y=f3f2, x=ChildAge, color=tran)
) + 
  scale_colour_manual(values=cbPalette) + theme_tufte()+
  geom_point()+
  geom_smooth(method=loess)
```




## scaled f2f3 by transcription
```{r}
ggplot(
  data=(acous %>% filter(
    #ChildID == unique(acous$ChildID)[22], 
    tran == c("w", "r")
  )), 
  aes(x=f3f2norm, color=tran)
) + 
  geom_density() + 
    scale_colour_manual(values=cbPalette) + theme_tufte()
```

# Adult Listener Ratings
```{r}
ggplot(data=rw.merge, aes(x=rating, color=transc))+geom_density(size=2)+
  scale_colour_manual(values=cbPalette) + theme_tufte()
```
The wr line is really flat ... which means the adult listeners ratings of wr transcriptions is ... suspect/unreliable. 

You have w's, and you have r's, and if you have even a  little bit of rhoticity, then people perceive it towards the /r/ side. /r/ bias. 

Need to check the method of transcription here, as there's some debate about how to group the transcriptions. People may perceive this is just R, W, or a half r, with some rhoticity but not full on. What does "wr" mean?

R makes sense. W makes sense. RW also makes sense. It's an R with a pinch of W. People aren't quite as sure with RW as the R, but it's pretty R like. The professional said this is neither r nor w but more r like, but the adult rates pushed them strongly towards R, but not fully so. 

## Target R/W and Produced R/W 

Adult rating by transcription by target word

$r is target w but sounds like r
$w is target r but sounds like w

```{r}
ggplot(data=(rw.merge %>% filter(!is.na(target))), aes(x=rating, color=transc))+geom_density(size=2)+
  scale_colour_manual(values=cbPalette) + theme_tufte()+
facet_wrap(~target)
```
No differences by target word. 



Is there a difference between r and $r? or w and $w? or r and $w vs. w and $r? 
```{r}
ggplot(data=(rw.merge %>% filter(!is.na(target), tran == c("$r", "r", "$w", "w"))), aes(x=rating, color=tran))+geom_density(size=2)+
  scale_colour_manual(values=cbPalette) + theme_tufte() 
```
This speaks to what Ben was investigating in his paper - is there a hidden difference in r productions based on target words? Maybe ever so slightly. 

Is there a difference in $r by age? 
```{r}
ggplot(data=(rw.merge %>% filter(!is.na(target), tran == c("$r", "$r:w"))), aes(x=ChildAge, y=rating, color=tran))+
  geom_point()+geom_smooth(method=lm)+
  scale_colour_manual(values=cbPalette) + theme_tufte() 
```

Is there a difference in $w by age? 
```{r}
ggplot(data=(rw.merge %>% filter(!is.na(target), tran == c("$w"))), aes(x=as.factor(ChildAge), y=rating, color=tran))+
#  geom_point()+geom_smooth(method=lm)+
  geom_violin()+
  scale_colour_manual(values=cbPalette) + theme_tufte() 
```

No. It seems age is only helpful when considering within subject development, not across all participants in a single time point study. 

## F3F2 distance and Rating

Is F3F2 distance distinct by transcription? 
```{r}
ggplot(data=(rw.merge %>% filter(!is.na(target), tran == c("r", "w"))), aes(x=f3f2, color=tran))+geom_density(size=2)+
  scale_colour_manual(values=cbPalette) + theme_tufte() 
```


```{r}
ggplot(data=(rw.merge %>% filter(!is.na(target), tran == c("$r", "r", "$w", "w"))), aes(x=f3f2, color=tran))+geom_density(size=2)+
  scale_colour_manual(values=cbPalette) + theme_tufte() 
```

Contour plots for f3f2 distance and rating by transcription. 

We're expecting ratings to change as f3f2 changes, i.e., some non-zero slope on the central axis.

```{r}
p1=ggplot(data=(rw.merge %>% filter(transc=="w")), aes(y=f3f2, x=rating)) + 
  labs(title="w") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(500,3000))

p2=ggplot(data=(rw.merge  %>% filter(transc=="r")), aes(y=f3f2, x=rating)) + 
  labs(title="r") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(500,3000))

p3=ggplot(data=(rw.merge  %>% filter(transc=="rw")), aes(y=f3f2, x=rating)) + 
  labs(title="rw") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(500,3000))

p4=ggplot(data=(rw.merge  %>% filter(transc=="wr")), aes(y=f3f2, x=rating)) + 
  labs(title="wr") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(500,3000))
ggarrange(p1,p2, p3, p4)
```

All these center axes look horizontal .... f3f2 isn't related to rating? Perhaps, with w and rw, it's at least related to variation. 


plot contours of ratings that are strictly r and w to look at f3f2 diff
```{r}
rw.merge %>% 
  filter(transc == "r" & rating < 0.125 | transc == "w" & rating > 0.875) %>% 
  ggplot() +
  aes(x = rating, y = f3f2) +
  geom_density_2d() +
  scale_y_continuous(limits=c(500,3000)) +
  scale_colour_manual(values=cbPalette) + theme_tufte() +
  facet_wrap(~transc, scales = "free")
```

```{r}
rw.merge %>% 
  filter(transc == "r" & rating < 0.125 | transc == "w" & rating > 0.875) %>% 
  ggplot() +
  aes(x = f3f2, color = transc) +
  geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() +
  scale_x_continuous(limits=c(500,3000))
```
Looks like very low f3f2 are more likely to be R, very high are W, in the middle, no difference. 

### F3 only

What about just for f3? 
```{r}
f1=ggplot(data=(rw.merge %>% filter(transc=="w")), aes(y=f3, x=rating)) + 
  labs(title="w") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(2000,5000))
f2=ggplot(data=(rw.merge %>% filter(transc=="r")), aes(y=f3, x=rating)) + 
  labs(title="r") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(2000,5000))
f3=ggplot(data=(rw.merge %>% filter(transc=="rw")), aes(y=f3, x=rating)) + 
  labs(title="rw") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(2000,5000))
f4=ggplot(data=(rw.merge %>% filter(transc=="wr")), aes(y=f3, x=rating)) + 
  labs(title="wr") + 
  geom_density_2d() + 
  scale_y_continuous(limits=c(2000,5000))
ggarrange(f1, f2, f3, f4)
```
F3 distances within a transcription are not related to the rating. But, super low f3 is more likely to be r than the others. 


What about f distances by transcription? 
```{r}
q1=ggplot(data=(rw.merge %>% filter(!is.na(transc))), aes(x=f3, color=transc))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() 
q2=ggplot(data=(rw.merge %>% filter(!is.na(transc))), aes(x=f2, color=transc))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() +
    scale_x_continuous(limits=c(500,4000))
q3=ggplot(data=(rw.merge %>% filter(!is.na(transc))), aes(x=f1, color=transc))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte()  +
    scale_x_continuous(limits=c(0,2000))
q4=ggplot(data=(rw.merge %>% filter(!is.na(transc))), aes(x=f3f2, color=transc))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() 
ggarrange(q1, q2, q3, q4)
```
No clear differences.... Maybe some slight ones, like wr for f3f2. Or w for f2. Or r for f3. 

### Vowel?

Front/Back Vowel influence on F2? 
```{r}
ggplot(data=(rw.merge %>% filter(Front.back==c("back","front"), !is.na(transc))), aes(x=f2, color=Front.back))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() +
  facet_wrap(~transc)+
  scale_x_continuous(limits=c(0,4000))

```
Says something about how the r and w is produced independently of the vowel. 

Vowel letter influence on F3F2, for R, back vowels? 
```{r}
ggplot(data=(rw.merge %>% filter(transc=="r", Letter!="ei", Front.back=="back")), aes(x=f3f2, color=Letter))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() +
  facet_wrap(~transc)+
  scale_x_continuous(limits=c(0,3000))
```

Vowel letter influence on F3F2, for R, front vowels? 
```{r}
ggplot(data=(rw.merge %>% filter(transc=="r", Letter!="ei", Front.back=="front")), aes(x=f3f2, color=Letter))+geom_density() +
  scale_colour_manual(values=cbPalette) + theme_tufte() +
  facet_wrap(~transc)+
  scale_x_continuous(limits=c(0,3000))
```








# w production and R production
```{r}
rw.parcoord.transc.r <- rw.merge %>% filter(transc=="r") %>% group_by(subject) %>% summarise(mean.transc.r = mean(rating))
rw.parcoord.transc.w <- rw.merge %>% filter(transc=="w") %>% group_by(subject) %>% summarise(mean.transc.w = mean(rating))
rw.parcoord.transc.rw <- rw.merge %>% filter(transc=="rw") %>% group_by(subject) %>% summarise(mean.transc.rw = mean(rating))
rw.parcoord.transc.wr <- rw.merge %>% filter(transc=="wr") %>% group_by(subject) %>% summarise(mean.transc.wr = mean(rating))
rw.parcoord.targ.r <- rw.merge %>% filter(target=="r") %>% group_by(subject) %>% summarise(mean.targ.r = mean(rating))
rw.parcoord.targ.w <- rw.merge %>% filter(target=="w") %>% group_by(subject) %>% summarise(mean.targ.w = mean(rating))

rw.pc <-
merge(
merge(
merge(
merge(
merge(
  rw.parcoord.targ.r, 
  rw.parcoord.targ.w, by="subject"), 
  rw.parcoord.transc.r, by="subject"),
  rw.parcoord.transc.wr, by="subject"),
  rw.parcoord.transc.w, by="subject"),
  rw.parcoord.transc.rw, by="subject")

rw.pc2 <- rw.pc %>% mutate(goodR = as.factor((mean.targ.r<.3)*1))
ggparcoord(rw.pc2,
    columns = c(3,2,4,7,5,6), groupColumn = 8, scale="globalminmax"
    ) +
  theme_tufte()+scale_colour_manual(values=c("red", "blue"))
```

```{r}
rw.pc.w <- rw.pc %>% mutate(goodW = as.factor(1-(mean.targ.w<.7)*1))
ggparcoord(rw.pc.w,
    columns = c(3,2,4,7,5,6), groupColumn = 8, scale="globalminmax"
    ) +
  theme_tufte()+scale_colour_manual(values=c("red", "blue"))
```
```{r}
ggplot(data=rw.pc, aes(x=mean.targ.w, y=mean.targ.r))+geom_point()+
  theme_tufte()+
  scale_x_continuous(limits=c(0,1))+  scale_y_continuous(limits=c(0,1))
```

```{r}
ggplot(data=rw.pc, aes(x=mean.transc.w, y=mean.transc.rw))+geom_point()+
  theme_tufte()+
  scale_x_continuous(limits=c(0,1))+  scale_y_continuous(limits=c(0,1))
```
